Current Models

### 2019-03-20 models.Rmd
more website edits
Website updates (#319)
* tidying up the website markdowns

* more website edits

* more website updates

* website edits

* Update report.Rmd

* Update about.md

* Update about.md

We currently analyze and forecast rodent data at Portal using four models:

@@

ESSS (Exponential Smoothing State Space) is a flexible exponential smoothing 
state space model [@Hyndman2008] fit to the data at the composite (full site
and just control plots) spatial level and the composite (community) ecological
level. The model is selected and fitted using the `ets` and `forecast` 
functions in the **forecast** package [@Hyndman2017] with the 
`allow.multiplicative.trend` argument set to `TRUE` and the `forecast_ess`
function in our **portalPredictions** package [@PortalPredictions]. 
Models fit using `ets` implement what is known as the "innovations" approach 
to state space modeling, which assumes a single source of noise that is 
equivalent for the process and observation errors [@Hyndman2008].

@@

AutoArima (Automatic Auto-Regressive Integrated Moving Average) is a 
flexible Auto-Regressive Integrated Moving Average (ARIMA) model fit to the 
data at the composite (full site and just control plots) spatial level
and the composite (community) ecological level. The model is selected and 
fitted using the `auto.arima` and `forecast` functions in the **forecast**
package [@Hyndman2013; @Hyndman2017] and the `forecast_autoarima` function 
in our **portalPredictions** package [@PortalPredictions].

@@

nbGARCH (Negative Binomial Auto-Regressive Conditional Heteroskedasticity) 
is a generalized autoregressive conditional heteroskedasticity (GARCH) 
model with overdispersion (*i.e.*, a negative binomial response) fit to the 
data at the composite (full site and just control plots) spatial level and both 
the composite (community) and the articulated (species) ecological levels. The 
model for each species and the community total is selected and fitted using the
`tsglm` function in the **tscount** package [@Liboschik2017a] and the 
`forecast_nbgarch` function in our **portalPredictions** package 
[@PortalPredictions]

@@

@@

pevGARCH (Poisson Environmental Variable Auto-Regressive Conditional
Heteroskedasticity) is a generalized autoregressive conditional
heteroskedasticity (GARCH) model fit to the data at the composite (full site 
and just control plots) spatial level and both the composite (community) and 
the articulated (species) ecological levels. The response variable is Poisson, 
and a variety of environmental variables are considered as covariates. The 
model for each species is selected and fitted using the `tsglm` function in the
**tscount** package [@Liboschik2017a] and the `forecast_pevgarch` function in
our **portalPredictions** package [@PortalPredictions].

>>>>

We currently analyze and forecast rodent data at Portal using five models:

<br>

@@

ESSS (Exponential Smoothing State Space) is a flexible exponential smoothing 
state space model [@Hyndman2008] fit to the data at the composite (full site
and just control plots) spatial level and both the composite (community) and 
the articulated (species) ecological levels. The model is selected and fitted 
using the `ets` and `forecast` functions in the **forecast** package 
[@Hyndman2017] with the `allow.multiplicative.trend` argument set to `TRUE` 
and the `ESSS` function in our 
[**portalcasting** package](https://github.com/weecology/portalcasting). 
Models fit using `ets` implement what is known as the "innovations" approach 
to state space modeling, which assumes a single source of noise that is 
equivalent for the process and observation errors [@Hyndman2008].

@@

AutoArima (Automatic Auto-Regressive Integrated Moving Average) is a 
flexible Auto-Regressive Integrated Moving Average (ARIMA) model fit to the 
data at the composite (full site and just control plots) spatial level
and both the composite (community) and the articulated (species) ecological 
levels. The model is selected and fitted using the `auto.arima` and `forecast`
 functions in the **forecast** package [@Hyndman2013; @Hyndman2017] and the 
`AutoArima` function in our 
[**portalcasting** package](https://github.com/weecology/portalcasting).

@@

nbGARCH (Negative Binomial Auto-Regressive Conditional Heteroskedasticity) 
is a generalized autoregressive conditional heteroskedasticity (GARCH) 
model with overdispersion (*i.e.*, a negative binomial response) fit to the 
data at the composite (full site and just control plots) spatial level and both 
the composite (community) and the articulated (species) ecological levels. The 
model for each species and the community total is selected and fitted using the
`tsglm` function in the **tscount** package [@Liboschik2017a] and the 
`nbGARCH` function in our
[**portalcasting** package](https://github.com/weecology/portalcasting).

@@

<br>

## nbsGARCH

nbsGARCH (Negative Binomial Seasonal Auto-Regressive Conditional 
Heteroskedasticity) is a generalized autoregressive conditional 
heteroskedasticity (GARCH) model with overdispersion (*i.e.*, a negative 
binomial response) with seasonal predictors modeled using two Fourier 
series terms (sin and cos of the fraction of the year) fit to the 
data at the composite (full site and just control plots) spatial level and both 
the composite (community) and the articulated (species) ecological levels. The 
model for each species and the community total is selected and fitted using the
`tsglm` function in the **tscount** package [@Liboschik2017a] and the 
`nbsGARCH` function in our
[**portalcasting** package](https://github.com/weecology/portalcasting).

GARCH models are generalized ARMA models and are defined according to their 
link function, response distribution, and two model structure parameters: the 
number of autoregressive terms (p) and the order of the moving average (q), 
and are represented as GARCH(p, q) [@Liboschik2017a]. The nbsGARCH model is fit
using the log link and a negative binomial response (modeled as an
over-dispersed Poisson), as well as with p = 1 (first-order autoregression)
and q = 12 (approximately yearly moving average).

<br>

@@

pevGARCH (Poisson Environmental Variable Auto-Regressive Conditional
Heteroskedasticity) is a generalized autoregressive conditional
heteroskedasticity (GARCH) model fit to the data at the composite (full site 
and just control plots) spatial level and both the composite (community) and 
the articulated (species) ecological levels. The response variable is Poisson, 
and a variety of environmental variables are considered as covariates. The 
model for each species is selected and fitted using the `tsglm` function in the
**tscount** package [@Liboschik2017a] and the `pevGARCH` function in
our [**portalcasting** package](https://github.com/weecology/portalcasting).
